Formalization of Forcing in Isabelle/ZF \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2020-01-25 IJCAR2020/conclusion.tex
pongo recuadro inline con un poco más

We consider that the formalization of the definition of $\forceisa$
and its recursive characterization of forcing for atomic formulas is
a turning point
in our project; the reason for this is that all further
developments will not involve such a daunting metamathematical
component. Even the proofs of the Fundamental Theorems of Forcing
turned out to follow rather smoothly after this initial setup was
ready, the only complicated affair being to show that various dense sets belong
to $M$. 
\begin{framed}
  Actually, this is a point to taken care of: the machinery of
  absoluteness and relativization must be developed for every new
  concept that is introduced.
\end{framed}

>>>>

We consider that the formalization of the definition of $\forceisa$
and its recursive characterization of forcing for atomic formulas is
a turning point
in our project; the reason for this is that all further
developments will not involve such a daunting metamathematical
component. Even the proofs of the Fundamental Theorems of Forcing
turned out to follow rather smoothly after this initial setup was
ready, the only complicated affair being to show that various dense sets belong
to $M$. 
%
Actually, this is a point to be taken care of: For every new
concept that is introduced, some lemmas concerning 
relativization and closure must be proved to be able to synthesize its
internal definition. Further automation must be developed for this
purpose.
